---
title: "Comparing manual vs automatically placed patches"
author: "Vera Weisbecker and Emma Sherratt"
date: "24 February 2020"
output: html_document
---



```{r}

#Code that is preliminary analysis but not dedicated to reading in (e.g. replicability analysis) is commented out to expedite read in

# Load libraries and version the code was last run on
library(rgl) #version 0.100.54
library(geomorph) #version 3.3.1
library(abind) #version 1.4-5
library(plyr) #version 1.8.6
library(vegan) #version 2.5-6
library(ape) #version 5.4-1
library(geiger) #version 2.0.7
library(nlme) #version 2.1-149
library(caper) #version 1.0.1
library(Morpho) #version 2.8


```


# Read in MANUAL patch placements

```{r }
# read in all others and set parameters


iter = 1000 # setting number of iterations for all of the phylogenetic analyses

# Read in classifiers and specimen volumes data 
ind.SpecData <- read.csv("../Data/Raw/Specimen_data.csv", header=T, row.names = 1 )

#For the sake of having everything in one spot, a lot of additional acquisition notes are in this file. Paring it down to what is required for this analysis
ind.SpecData <- ind.SpecData[,c(1:13, 19:20)]


####~~~~~~~~~~~~~~loading 3d coordinate files~~~~~~~~~~~~~~~~~####

#FIRST REPLICATE DATASET
filelist <- list.files(path= "../Data/Raw/Replicates_1", pattern = "*.txt")
names <- gsub (".txt", "", filelist) # extracts names of specimens from the file name
filelist <- paste("../Data/Raw/Replicates_1/", filelist, sep="") # rename with path
coords <- NULL # make empty object that will be filled with 3D array of coordinate data
for (i in 1:length(filelist)){
  temp <- read.morphologika(filelist[i])
  k <- dim(temp)[1]
  coords <- rbind(coords, two.d.array(temp)) }
rep.1 <- arrayspecs(coords, k, 3)
dimnames(rep.1)[[3]] <- names
remove(i, filelist, names, k, coords, temp) # clean up environment

##SECOND REPLICATE DATASET
filelist <- list.files(path= "../Data/Raw/Replicates_2", pattern = "*.txt")
names <- gsub (".txt", "", filelist) # extracts names of specimens from the file name
filelist <- paste("../Data/Raw/Replicates_2/", filelist, sep="") # rename with path
coords <- NULL # make empty object that will be filled with 3D array of coordinate data
for (i in 1:length(filelist)){
  temp <- read.morphologika(filelist[i])
  k <- dim(temp)[1]
  coords <- rbind(coords, two.d.array(temp)) }
rep.2 <- arrayspecs(coords, k, 3)
dimnames(rep.2)[[3]] <- names
remove(i, filelist, names, k, coords, temp) # clean up environment

Manual_raw <- abind(rep.1,rep.2) # abind concatenates (stacks) two arrays
remove(rep.1,rep.2) # clean up environment
```

# Read in AUTOMATIC patch placements

```{r}



load("../Data/Processed/Automatic_Patch_Placement_raw.rda")


```


#GPAs

```{r}

## define sliding semilandmarks on curves (curveslide), and patches (surfslide); these are DIFFERENT in automatic patch placement! 
curveslide_manual=as.matrix(read.csv("../Data/Raw/curveslide_manual.csv", header=T))
surfslide_manual=as.matrix(read.csv("../Data/Raw/surfslide_manual.csv",header=T))

## define sliding semilandmarks on curves (curveslide), and patches (surfslide)
curveslide_automatic=as.matrix(read.csv("../Data/Raw/curveslide_automatic.csv", header=T))
surfslide_automatic=as.matrix(read.csv("../Data/Raw/surfslide_automatic.csv",header=T))

### GPA of both sets of replicates, with sliding
Manual_GPA <- gpagen(Manual_raw, surfaces = surfslide_manual, curves = curveslide_manual)
Automatic_GPA <- gpagen(Automatic_Raw, surfaces = surfslide_automatic, curves = curveslide_automatic)

#test if all worked, commented out here for speed of running script

# plot3d(Manual_GPA$coords[,,5], asp=FALSE)
# 
# plot3d(Automatic_GPA$coords[,,5], asp=FALSE)

#Quick check to see if the csizes are equivalent - an important check because some plys were exported from Mimics too large for technical reasons (refer to Patch Placement code)
#plot(Manual_GPA$Csize~ Automatic_GPA$Csize)
#text(Manual_GPA$Csize~ Automatic_GPA$Csize, labels=(dimnames(Manual_GPA$coords)[[3]]))

#remove objects no longer needed
remove(curveslide_manual, surfslide_manual, curveslide_automatic, surfslide_automatic)

```


#compare replication error - commented out here to run script quicker

```{r}

n=length(dimnames(Manual_GPA$coords[,,])[[3]])/2
individuals <- dimnames(Manual_GPA$coords)[[3]] # name for each pair shouldbe same, so not _1, _2
replicate <- c(rep(1, n), rep(2, n)) # creates a vector designating specimens to a replicate


#Repeatability for Manual coordinates
rep.er <- procD.lm(Manual_GPA$coords ~ factor(individuals))
((rep.er$aov.table$MS[1] - rep.er$aov.table$MS[2])/2) / (rep.er$aov.table$MS[2] + ((rep.er$aov.table$MS[1] - rep.er$aov.table$MS[2])/2))


# repeatability for Automatically placed coordinates
rep.er <- procD.lm(Automatic_GPA$coords ~ factor(individuals))
((rep.er$aov.table$MS[1] - rep.er$aov.table$MS[2])/2) / (rep.er$aov.table$MS[2] + ((rep.er$aov.table$MS[1] - rep.er$aov.table$MS[2])/2))

remove(n, individuals,rep.er, replicate)

```

#Compare how replication error differences may affect the main variation:

```{r}
# par(mfrow=c(2,1))
# 
 PCA_Manual <- gm.prcomp(Manual_GPA$coords)
# 
# PCA_summary <- summary(PCA_Manual)
# 
# #Ugh I miss the quick simplicity of PlotTangetSpace!
# plot(PCA_Manual$x[,2] ~ PCA_Manual$x[,1], pch = 19)
# text( PCA_Manual$x[,2] ~ PCA_Manual$x[,1],  labels = dimnames (Manual_GPA$coords)[[3]])
# 
PCA_Automatic <- gm.prcomp(Automatic_GPA$coords)
# 
# plot(PCA_Automatic$x[,2] ~ PCA_Automatic$x[,1], pch = 19)
# text( PCA_Automatic$x[,2] ~ PCA_Automatic$x[,1],  labels = dimnames (Automatic_GPA$coords)[[3]])
# 
# #The morphospaces are very similar; this can be tested across all PCAs using a Mantel test in the Vegan package:
# 
 Manual_dist <- dist(PCA_Manual$x) ; 
 Automatic_dist <- dist(PCA_Automatic$x)
 
 #The matrix correlation is 0.97, so the protocols seem equivalent.
 mantel(Manual_dist,Automatic_dist,method="pearson", permutations=9999) 
 
 remove (Manual_dist, Automatic_dist)


```

# Make a figure to show that the estimated main variation as per PC1 is near-identical
```{r}

usrMat.dorsal<- as.matrix (read.csv("../Figures/usrMat.dorsal.csv"))

open3d( windowRect=c(0,0,1200,1200), userMatrix=usrMat.dorsal)

mfrow3d(nr=1, nc=2)

plotRefToTarget(PCA_Automatic$shapes$shapes.comp1$min, mshape(Automatic_GPA$coords), method="vector" )
plotRefToTarget(PCA_Manual$shapes$shapes.comp1$min, mshape(Manual_GPA$coords), method="vector" )

rgl.snapshot("../Figures/Supp_PC1_comp.png", fmt = "png", top=TRUE )

remove(Manual_dist,Automatic_dist, PCA_Manual, PCA_Automatic)

```


# (not reported in ms but just out of interest - P. ingrami such an outlier in the automatic placement mainly because it better captures the flatness and ventral concavity.
```{r}
# 
# rgl.open()
# bg3d("white")
# 
# plotRefToTarget(mshape(Manual_GPA$coords), Manual_GPA$coords[,,141], method="vector" )
# 
# 
# rgl.open()
# bg3d("white")
# #The automatic placement seems "truer" to the extreme width and flatness, but not much so. 
# plotRefToTarget(mshape(Automatic_GPA$coords), Automatic_GPA$coords[,,141], method="vector" )

```

#Averaging replicates and removing symmetry

```{r}

#landpair files 

landpairs_manual <- as.matrix(read.csv("../Data/Raw/Landpairs_manual.csv", header = F))
landpairs_automatic <- as.matrix(read.csv("../Data/Raw/Landpairs_automatic.csv", header = F))


#MANUAL Patch

# shape data
ind.coords_Manual <- aggregate(two.d.array(Manual_GPA$coords) ~ dimnames(Manual_GPA$coords)[[3]], FUN=mean)[,-1]
rownames(ind.coords_Manual) <- unique(dimnames(Manual_GPA$coords)[[3]])
ind.coords_Manual <- arrayspecs(ind.coords_Manual, p=ncol(ind.coords_Manual)/3, k=3)
# centroid size data
ind.csize_Manual <- as.vector(aggregate(Manual_GPA$Csize ~ dimnames(Manual_GPA$coords)[[3]], FUN=mean)[,-1])
names(ind.csize_Manual) <- unique(dimnames(Manual_GPA$coords)[[3]])
# ind.csize_Manual - These are the individuals Csize data for plotting specimen allometry graphs

### Bilateral symmetry GPA on averaged specimens

GPASymm <- bilat.symmetry(ind.coords_Manual, ind=dimnames(ind.coords_Manual)[[3]], side=NULL, replicate=NULL, object.sym=TRUE, land.pairs =landpairs_manual)
dimnames(GPASymm$symm.shape)[[3]] <- dimnames(ind.coords_Manual)[[3]] # rename

# ind.coords_Manual - These are the individuals shape data for plotting specimen graphs
ind.coords_Manual <- GPASymm$symm.shape # replace with new symm.shape

#AUTOMATIC Patch - same code


# shape data
ind.coords_Automatic <- aggregate(two.d.array(Automatic_GPA$coords) ~ dimnames(Automatic_GPA$coords)[[3]], FUN=mean)[,-1]
rownames(ind.coords_Automatic) <- unique(dimnames(Automatic_GPA$coords)[[3]])
ind.coords_Automatic <- arrayspecs(ind.coords_Automatic, p=ncol(ind.coords_Automatic)/3, k=3)

ind.csize_Automatic <- as.vector(aggregate(Automatic_GPA$Csize ~ dimnames(Automatic_GPA$coords)[[3]], FUN=mean)[,-1])
names(ind.csize_Automatic) <- unique(dimnames(Automatic_GPA$coords)[[3]])


GPASymm <- bilat.symmetry(ind.coords_Automatic, ind=dimnames(ind.coords_Automatic)[[3]], side=NULL, replicate=NULL, object.sym=TRUE, land.pairs =landpairs_automatic)
dimnames(GPASymm$symm.shape)[[3]] <- dimnames(ind.coords_Automatic)[[3]] # rename

ind.coords_Automatic <- GPASymm$symm.shape # replace with new symm.shape

remove(landpairs_automatic,landpairs_manual)

```

# This is to just briefly check that the symmetry was removed successfully (i.e. the landpair files were correct); commented out for speedily running the code.

```{r}
# 
# rgl.open()
# bg3d("white")
# 
# plot3d(ind.coords_Manual[,,5], asp=FALSE)
# text3d ( ind.coords_Manual[,,5], texts =  c(1:71))
# 
# 
# 
# rgl.open()
# bg3d("white")
# 
# plot3d(ind.coords_Automatic[,,5], asp=FALSE)
# text3d ( ind.coords_Automatic[,,5], texts =  c(1:71))
# 
# 
# rgl.open()
# bg3d("white")
# 
# plotRefToTarget(mshape(ind.coords_Manual), ind.coords_Manual[,,5], method="vector" )
# 
# 
# rgl.open()
# bg3d("white")
# 
# plotRefToTarget(mshape(ind.coords_Automatic), ind.coords_Automatic[,,5], method="vector" )



```




#Averaging by species, this chunk treats automatic and manual at the same time

```{r}

# Make a vector of species labels from the specimen names
species <- matrix(unlist(strsplit(dimnames(ind.coords_Manual)[[3]], "_")), ncol=2, byrow = T)[,1]

# Reorder the classifier file to match this up with shape data
ind.SpecData <- ind.SpecData[match(dimnames(ind.coords_Manual)[[3]],rownames(ind.SpecData)),]

#check that re-ordering worked for automatic and manual
match(dimnames(ind.coords_Manual)[[3]],rownames(ind.SpecData))==c(1:84)
match(dimnames(ind.coords_Automatic)[[3]],rownames(ind.SpecData))==c(1:84)

# Add a column of species labels from the specimen names
ind.SpecData$Abbreviation <- species

#Add a column for just macropodoids vs the rest
ind.SpecData$Macropod <- NA

## Filling it with the two different factors
ind.SpecData$Macropod=as.factor(ifelse(as.character(ind.SpecData$Cladename) == "Macropodoid", "yes", "no"))


# Average classifiers and volumes data by species
SpecData <- ddply(ind.SpecData, .(Species), summarize, Clade_major=unique(Clade_major),Cladename=unique(Cladename),Locomotion=unique(Locomotion), Fossil=unique(Fossil), Macropod = unique(Macropod), Bodymass_g=mean(Bodymass_g),Brain_mm3=mean(Brain_mm3), OB=mean(OB), Cerebrum=mean(Cerebrum), Cerebellum=mean(Cerebellum), Medulla=mean(Medulla), Grey_Matter_LSR = unique(Iso_and_periallo_Cx_GM_vol_Jyothilakshmi), Neocortex_LSR = unique(Neocortex_Pirlot), Abbreviation=unique(Abbreviation))

SpecData <- SpecData[order(SpecData$Abbreviation),]# Alphabetical by Gsp abbreviation

# Averaging shape data by species

#MANUAL

coords_Manual <- (aggregate(two.d.array(ind.coords_Manual) ~ ind.SpecData$Abbreviation, FUN=mean))[,-1]
rownames(coords_Manual) <- unique(ind.SpecData$Abbreviation)
coords_Manual <- arrayspecs(coords_Manual,p=dim(ind.coords_Manual)[1],k=3) # 3D array

#AUTOMATIC

coords_Automatic <- (aggregate(two.d.array(ind.coords_Automatic) ~ ind.SpecData$Abbreviation, FUN=mean))[,-1]
rownames(coords_Automatic) <- unique(ind.SpecData$Abbreviation)
coords_Automatic <- arrayspecs(coords_Automatic,p=dim(ind.coords_Automatic)[1],k=3) # 3D array



#Checking a distinctive shape (P. ingrami) to ensure all is well
dimnames(coords_Automatic)[[3]][38]
plot3d(coords_Automatic[,,38], asp=FALSE)
text3d(coords_Automatic[,,38], texts=c(1:71))

# Averaging centroid size by species
#MANUAL
Csize_Manual<- as.vector(aggregate(as.matrix(ind.csize_Manual ) ~ ind.SpecData$Abbreviation, FUN=mean)[,-1])
names(Csize_Manual)<-unique(ind.SpecData$Abbreviation)

#AUTOMATIC

Csize_Automatic<- as.vector(aggregate(as.matrix(ind.csize_Automatic ) ~ ind.SpecData$Abbreviation, FUN=mean)[,-1])
names(Csize_Automatic)<-unique(ind.SpecData$Abbreviation)


### Reorder SpecData to match the shape and csize datasets - this needs doing only once
dimnames(coords_Automatic)[[3]]==dimnames(coords_Manual)[[3]]


SpecData <- SpecData[match(SpecData$Abbreviation, dimnames(coords_Automatic)[[3]]),] 

#Check order
match(SpecData$Abbreviation, dimnames(coords_Manual)[[3]]) == c(1:57)

rownames(SpecData) <- SpecData$Species

### Rename to match up with tree below
dimnames(coords_Manual)[[3]] <- names(Csize_Manual) <- SpecData$Species
dimnames(coords_Automatic)[[3]] <- names(Csize_Automatic) <- SpecData$Species

```


#Now read in the phylogeny

```{r}

# Read in trees
# ALREADY RUN and "paleotree_dated_Yalk_resolved.tre" in folder so ignore indented below
# Tree1=read.nexus("Phylogenies/paleotree_dated_Yalk1.tre")
# Tree2=read.nexus("Phylogenies/paleotree_dated_Yalk2.tre")
# Tree3=read.nexus("Phylogenies/paleotree_dated_Yalk3.tre")
# write.nexus(list(Tree1=Tree1,Tree2=Tree2,Tree3=Tree3), file="Phylogenies/paleotree_dated_Yalk_unresolved.tre")
## There are 3 trees, with "Yalkaparidon_coheni" in different positions
# Tree 1 has 9 polytomies
# Tree 2 has 10 polytomies
# Tree 3 has 9 polytomies
# Therefore it is necessary to pseudoresolve the trees prior to analysis.
# Randomly resolve polytomies to zero branch lengths
# treelist <- list(Tree1 = multi2di(Tree1, random = TRUE), 
#                  Tree2 = multi2di(Tree2, random = TRUE), 
#                  Tree3 = multi2di(Tree3, random = TRUE)) 
# # run multi2di on each tree and concatenate into a list
# write.nexus(treelist, file="Phylogenies/paleotree_dated_Yalk_resolved.tre")
# remove(treelist,Tree1,Tree2,Tree3)

# Read in New Trees file
# a nexus file of the three topologies, which have the polytomies randomly resolved with 0 length branches
trees <- read.nexus("../Data/Raw/Phylogenies/paleotree_dated_Yalk_resolved.tre") 

# reorders to tip labels; all trees have same tip order (see above)
coords_Manual <- coords_Manual[,,trees$Tree1$tip.label] 
coords_Automatic <- coords_Automatic[,,trees$Tree1$tip.label] 

# reorders to tip labels
Csize_Manual <- Csize_Manual[trees$Tree1$tip.label] 
Csize_Automatic <- Csize_Automatic[trees$Tree1$tip.label] 

SpecData <- SpecData[trees$Tree1$tip.label,] # reorders to tip labels

#Test that dimnames and tree tip labels are ordered correctly
match(dimnames(coords_Automatic)[[3]], trees$Tree1$tip.label) == c(1:length(trees$Tree1$tip.label))
match(dimnames(coords_Manual)[[3]], trees$Tree1$tip.label) == c(1:length(trees$Tree1$tip.label))


#Make gdf data frame 

gdf_manual <- geomorph.data.frame(coords_Manual=coords_Manual, Csize_Manual=Csize_Manual, SpecData=SpecData)

gdf_manual$coords <- gdf_manual$coords_Manual
gdf_manual$Csize <- gdf_manual$Csize_Manual

attributes(gdf_manual)

gdf_manual <- gdf_manual[-c(13,14)]
attributes(gdf_manual)



gdf_automatic <- geomorph.data.frame(coords=coords_Automatic, Csize=Csize_Automatic, SpecData=SpecData)


attributes(gdf_automatic)


```

#Save the workspace - it was cleaned up as I went along to contain what is necessary for the next analyses

```{r}
save.image(file = "../Data/Processed/brain_data_processed.rda")


```
