---
title: "02_1_Marsbrain_Analyses"
author: "Vera Weisbecker & Emma Sherratt"
date: "30/11/2020"
output: html_document
---

---
title: "02_Marsbrain_analyses"
author: "Vera Weisbecker & Emma Sherratt"
date: "25/11/2020"
output: html_document
---

# Load libraries
```{r}

# Load libraries and version the code was last run on

library(rgl) # version 0.100.54
library(geomorph) #  version 3.3.1
library(abind) # version 1.4.5
library(plyr) # version 1.8.6
library(vegan) # version 2.5-6
library(ape) # version 5.4-1
library(geiger)# version 2.0.7
library(nlme) # version 3.1-149
library(caper) # version 1.0.1
library(phytools) # version 0.7-10

load("../Data/Processed/brain_data_processed.rda")

```

#Adding a little function to summarize ANOVA results over multiple trees (note that this is only semi-automated - some df numbers need to be entered manually into the published table)

```{r}

mean.tree.results <- function (Resultslist){
   
    c (
   
  round(Resultslist$Tree1$aov.table$`Df`[2]),
      
  round(mean(Resultslist$Tree1$aov.table$F[1],
       Resultslist$Tree2$aov.table$F[1],
       Resultslist$Tree3$aov.table$F[1]), 
       digits = 3),# average F
 
  
 round(mean(Resultslist$Tree1$aov.table$Rsq[1],
       Resultslist$Tree2$aov.table$Rsq[1],
       Resultslist$Tree3$aov.table$Rsq[1]),
 digits = 2), # average Rsq
  
 round( mean(Resultslist$Tree1$aov.table$`Pr(>F)`[1],
       Resultslist$Tree2$aov.table$`Pr(>F)`[1],
       Resultslist$Tree3$aov.table$`Pr(>F)`[1]),
  digits = 3) # average p 
                )
   
}

ANOVA_results <- data.frame( "Df"= integer(0), "F"= integer(0), "Rsq"= integer(0), "p"= integer(0))



```

# PCA & Summary Stats

```{r}

PCA <- gm.prcomp(gdf_automatic$coords)

plot(PCA$x[,2] ~ PCA$x[,1], pch = 19)
text( PCA$x[,2] ~ PCA$x[,1],  label = dimnames(gdf_automatic$coords)[[3]])

#Checking that the deformations are OK even in the outlier P. ingrami  - it's definitely the correct shape.
which(dimnames(gdf_automatic$coords)[[3]]=="Planigale_ingrami")
plotRefToTarget(gdf_automatic$coords[,,43], mshape(gdf_automatic$coords),  method = "vector")

#Brief check of main PC shape variation using lollipop graphs

gridpars <- gridPar(pt.size=0.5, pt.bg="purple")
plotRefToTarget(mshape(gdf_automatic$coords), PCA$shapes$shapes.comp1$max, method="vector",gridPars = gridpars)
plotRefToTarget(mshape(gdf_automatic$coords), PCA$shapes$shapes.comp2$max, method="vector", gridPars = gridpars)

plot(PCA$x[,2] ~ PCA$x[,1], pch = 19)
text( PCA$x[,2] ~ PCA$x[,1],  labels=gdf_automatic$SpecData.Abbreviation)


```


#All subsequent analyses do the same thing over each tree 

# phylogenetic signal
```{r}

physig_shape  <- list(Tree1=NULL, Tree2=NULL,Tree3=NULL)
physig_size   <- list(Tree1=NULL, Tree2=NULL,Tree3=NULL)

for (i in 1:length(physig_shape)){
  physig_shape[[i]] <- physignal(gdf_automatic$coords, trees[[i]])
  physig_size[[i]] <- physignal(log(gdf_automatic$Csize), trees[[i]])
}

#extract mean phylogenetic signal and p-values over all three trees
mean(physig_shape$Tree1$phy.signal,physig_shape$Tree2$phy.signal, physig_shape$Tree3$phy.signal)
mean(physig_shape$Tree1$pvalue,physig_shape$Tree2$pvalue,physig_shape$Tree3$pvalue)

mean(physig_size$Tree1$phy.signal,physig_size$Tree2$phy.signal, physig_size$Tree3$phy.signal)
mean(physig_size$Tree1$pvalue,physig_size$Tree2$pvalue,physig_size$Tree3$pvalue)

```


#Allometry

```{r}

#first, just for later visualizing non-evolutionary allometry

# Is centroid size a good proxy of ecv?

Csize_brainvol <- summary(lm(log(gdf_automatic$Csize)~log(SpecData$Brain_mm3)))

Csize_brainvol_output <- c(Csize_brainvol$df[2], round(Csize_brainvol$fstatistic[1], 3), round(Csize_brainvol$r.squared, 3), round(Csize_brainvol$coefficients[2,4], 3))

ANOVA_results <- rbind(ANOVA_results, Csize_brainvol_output); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Csize_Brainvol"

#This is non-phylogenetic but required for figures
Allom <- procD.lm(coords ~ Csize,iter=iter, data = gdf_automatic, logsz = TRUE)

summary(Allom)

## Evolutionary Allometry - Using centroid size
  res.phy.allom.CS <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)
  for(i in 1:3){
     res.phy.allom.CS[[i]] <- procD.pgls(coords ~ log(Csize), trees[[i]],
                                     iter=iter, data=gdf_automatic)}

  #This line uses the mean tree results function entered at the top of this code to add a row of results, note the soft coding of the row re-name
  ANOVA_results <- rbind(ANOVA_results, mean.tree.results(res.phy.allom.CS)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_logCsize"


  ### Evolutionary Allometry - Using Body Mass (g)
  res.phy.allom.mass <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)
  for(i in 1:3){
    res.phy.allom.mass [[i]] <- procD.pgls(coords ~ log(SpecData.Bodymass_g),
                                       trees[[i]], iter=iter, data=gdf_automatic)}



  ANOVA_results <- rbind(ANOVA_results, mean.tree.results(res.phy.allom.mass)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_logBodyMass"



  ### Evolutionary Allometry -  relative brain size (EQ)

    # using the residuals of a regression of log(brain size) on log(body size)

  #make brain/body data frame and make sure it matches the tree
  brain_body <- as.data.frame(cbind(gdf_automatic$SpecData.Bodymass_g,gdf_automatic$SpecData.Brain_mm3))
  rownames(brain_body) <- gdf_automatic$SpecData.Species
  colnames(brain_body) <- c("Body", "Brain")
  trees$Tree1$tip.label == rownames(brain_body)

  #the below gls coefficient is close to the slope found for other marsupials ( Weisbecker & Goswami 2011), we are thus using it to obtain residuals for our specific dataset (rather than a larger published one)
  brain_body_gls <- (gls(log(Brain) ~ log(Body), correlation = corBrownian(1, phy=trees$Tree1), data=brain_body))

  summary (brain_body_gls)

  EQ <- residuals(brain_body_gls)
  names(EQ) <- rownames(gdf_automatic$SpecData.Species)
  res.phy.allom.EQ <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

    for(i in 1:3){
      res.phy.allom.EQ[[i]] <- procD.pgls(coords ~ EQ, trees[[i]],iter=iter, data = gdf_automatic)
    }

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(res.phy.allom.EQ)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Residuals"
```

# Vombatiforms and the Bandicoots are clearly particularly non-allometric. Removing them from allometry analyses

```{r}

  omit <- c(which(gdf_automatic$SpecData.Cladename == "Vombatiform"),
            which(gdf_automatic$SpecData.Cladename == "Peramelemorph")) # save which species to omit

  #Make new data frame. NOTE that the analysis seems to only work if the logging is done at the gdf_automatic building step, rather than later. See similar issue below with the volume dataset.

  gdf_automatic_VP<- geomorph.data.frame(coords=gdf_automatic$coords[,,-omit], Csize=log(gdf_automatic$Csize[-omit]),
                               bmass = gdf_automatic$SpecData.Bodymass_g[-omit])

  #It is not necessary to reduce the tree, the results remain identical
  # trees_VP <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)
  #
  #   for(i in 1:3){
  #     trees_VP[[i]] <- drop.tip(trees[[i]], omit)
  #
  #   }

      ### using Csize
     res_phy_allom_CS_VP <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

     #This is really frustrating - it seems that logging the data within the procD.pgls formula stops the function from working.


     for(i in 1:3){
       res_phy_allom_CS_VP[[i]] <- procD.pgls(coords ~ Csize,
                                     trees[[i]],
                                     iter=iter, data=gdf_automatic_VP)
       }


    ANOVA_results <- rbind(ANOVA_results, mean.tree.results(res_phy_allom_CS_VP)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_logCsize_VP"

    ### using Body Mass (g)
    res.phy.allom.mass_VP <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

    for(i in 1:3){
      res.phy.allom.mass_VP[[i]] <- procD.pgls(coords ~ log(bmass),
                                trees[[i]],
                                iter=iter, data=gdf_automatic_VP)
      }

     ANOVA_results <- rbind(ANOVA_results, mean.tree.results(res.phy.allom.mass_VP)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_logBody_mass_VP"





# Lastly, a pgls suggests that evolutionary allometry of PC1 is weak

PCallom <- as.data.frame(cbind(PCA$x[,1], log(gdf_automatic$Csize)))
colnames (PCallom) <- c("PC1", "Csize")

 PC1_allom_results <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

  for(i in 1:3){
     PC1_allom_results[[i]] <- procD.pgls(PC1 ~ Csize, trees[[i]],
                                     iter=iter, data=PCallom)}

 ANOVA_results <-  rbind(ANOVA_results, mean.tree.results(PC1_allom_results)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] = "PC1_vs_logCsize"

```

#Analyses of locomotor mode. I find it useful to always randomly spot-check that all the data are in the right order (e.g. is Acrobates really listed as "gliding")

```{r}

dimnames(coords_Automatic)[[3]]==rownames(SpecData)

#Note Csize is logged here at the start to avoid complications. Taking out unknown locomotion and Notoryctes which is the only fossorial species. Note that you need to take Notoryctes out first and then the unknown locomotor modes, the other way around causes an error message.

gdf_loco <- geomorph.data.frame(coords=coords_Automatic [,,which(SpecData$Species!= "Notoryctes_typhlops" & !is.na(SpecData$Locomotion))], Csize = log(Csize_Automatic[which(SpecData$Species!= "Notoryctes_typhlops" & !is.na(SpecData$Locomotion))]), Locomotion = SpecData$Locomotion[which(SpecData$Species!= "Notoryctes_typhlops" & !is.na(SpecData$Locomotion))])




#There are two species listed as "scansorial"; running these analyses with them classified as arboreal due to their additional climbing habit from a terrestrial ancestry
gdf_loco$Locomotion[which (gdf_loco$Locomotion == "Scansorial") ] <- rep("Arboreal", length (which (gdf_loco$Locomotion == "Scansorial")) )

# Not reported in ms but classifying the scansorial species as scansorial or as terrestrial  also doesn't do anything
#gdf_loco$Locomotion[which (gdf_loco$Locomotion == "Scansorial") ] <- rep("Terrestrial", length (which (gdf_loco$Locomotion == "Scansorial")) )



#First, making sure that locomotion analyses are not confounded by interactions on mean differences with centroid size. t First testing interactions

Loco_list_Csize_inter <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)
  for(i in 1:3){
     Loco_list_Csize_inter[[i]] <- procD.pgls(coords ~ Csize*Locomotion, trees[[i]], iter=iter, data=gdf_loco)
     }


# we need to separately enter the values into the table because the first row is Csize statistics, 2nd one is locomotor statistics, 3rd is interaction term

Loco_list_Csize_inter <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)
  for(i in 1:3){
     Loco_list_Csize_inter[[i]] <- procD.pgls(coords ~ Csize*Locomotion, trees[[i]], iter=iter, data=gdf_loco)
     }


ANOVA_results [length(rownames(ANOVA_results)),] <- c(

  Loco_list_Csize_inter$Tree1$aov.table[4,1],

  
 round( mean(Loco_list_Csize_inter$Tree1$aov.table$F[3],
       Loco_list_Csize_inter$Tree2$aov.table$F[3],
       Loco_list_Csize_inter$Tree3$aov.table$F[3]),
       digits=3), # average F


 round  (mean(Loco_list_Csize_inter$Tree1$aov.table$Rsq[3],
       Loco_list_Csize_inter$Tree2$aov.table$Rsq[3],
       Loco_list_Csize_inter$Tree3$aov.table$Rsq[3]), digits = 2), # average Rsq

  round( mean(Loco_list_Csize_inter$Tree1$aov.table$`Pr(>F)`[3],
       Loco_list_Csize_inter$Tree2$aov.table$`Pr(>F)`[3],
       Loco_list_Csize_inter$Tree3$aov.table$`Pr(>F)`[3]),
       digits=3)# average p
)

rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Locomotor_mode_Csize_inter1"



# Now testing for slope differences relative to csize

Loco_list_Csize_adj <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)
  for(i in 1:3){
     Loco_list_Csize_adj[[i]] <- procD.pgls(coords ~ Csize+Locomotion, trees[[i]], iter=iter, data=gdf_loco)
     }

#This time, second row position
ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Loco_list_Csize_adj)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Locomotor_mode_Csize_adj"

ANOVA_results [length(rownames(ANOVA_results)),] <- c(

   Loco_list_Csize_adj$Tree1$aov.table[4,1],

 round( mean(Loco_list_Csize_adj$Tree1$aov.table$F[2],
       Loco_list_Csize_adj$Tree2$aov.table$F[2],
       Loco_list_Csize_adj$Tree3$aov.table$F[2]),
       digits=3), # average F


 round  (mean(Loco_list_Csize_adj$Tree1$aov.table$Rsq[2],
       Loco_list_Csize_adj$Tree2$aov.table$Rsq[2],
       Loco_list_Csize_adj$Tree3$aov.table$Rsq[2]), digits = 2), # average Rsq

  round( mean(Loco_list_Csize_adj$Tree1$aov.table$`Pr(>F)`[2],
       Loco_list_Csize_adj$Tree2$aov.table$`Pr(>F)`[2],
       Loco_list_Csize_adj$Tree3$aov.table$`Pr(>F)`[2]),
       digits=3)# average p
)

rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Locomotor_mode_Csize_adj"


#Now just locomotion since Csize does not seem to contribute to the model

Loco_list <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)
  for(i in 1:3){
     Loco_list[[i]] <- procD.pgls(coords ~ Locomotion, trees[[i]], iter=iter, data=gdf_loco)
     }

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Loco_list)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Locomotor_mode"

```

#But there could be an issue with the very unusual mode of hopping within macropods. So comparing the shape of hopping vs. non-hopping Diprotodontians only (because of the reasonably distinctive shape of diprotodontian brains)

```{r}

gdf_hopping <- gdf_automatic


gdf_hopping <-
  geomorph.data.frame(
    coords = gdf_hopping$coords[, , which(SpecData$Clade_major == "Diprotodontia" &
                                           !is.na(SpecData$Locomotion))],
    Csize = log(gdf_hopping$Csize[which(SpecData$Clade_major == "Diprotodontia" &
                                  !is.na(SpecData$Locomotion))]),
    Locomotion = gdf_hopping$SpecData.Locomotion[which(SpecData$Clade_major == "Diprotodontia" &
                                       !is.na(SpecData$Locomotion))]
  )


gdf_hopping$Locomotion[which (gdf_hopping$Locomotion != "Hopping") ] <- rep("Non-hopping", length (which (gdf_hopping$Locomotion != "Hopping")) )



#No difference in shape between hoppers and non-hoppers

Loco_list_hop <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)
  for(i in 1:3){
     Loco_list_hop[[i]] <- procD.pgls(coords ~ Locomotion, trees[[i]], iter=iter, data=gdf_hopping)
     }


ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Loco_list_hop)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_hopping"



```

#What about gliders?

```{r}

gdf_gliding <- gdf_automatic


gdf_gliding <-
  geomorph.data.frame(
    coords = gdf_gliding$coords[, , which(SpecData$Clade_major == "Diprotodontia" &
                                           !is.na(SpecData$Locomotion))],
    Csize = log(gdf_gliding$Csize[which(SpecData$Clade_major == "Diprotodontia" &
                                  !is.na(SpecData$Locomotion))]),
    Locomotion = gdf_gliding$SpecData.Locomotion[which(SpecData$Clade_major == "Diprotodontia" &
                                       !is.na(SpecData$Locomotion))]
  )


gdf_gliding$Locomotion[which (gdf_gliding$Locomotion != "Gliding") ] <- rep("Non-gliding", length (which (gdf_gliding$Locomotion != "Gliding")) )



#No difference in shape between glidpers and non-glidpers

Loco_list_glid <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)
  for(i in 1:3){
     Loco_list_glid[[i]] <- procD.pgls(coords ~ Locomotion, trees[[i]], iter=iter, data=gdf_gliding)
     }


ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Loco_list_glid)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_gliding"



```
# Analyses including brain volume

```{r}
# PCA of the brain volume data - volumes for the 4 partitions
Volumes<- data.frame(OB=SpecData$OB, CBR=SpecData$Cerebrum,
                     CBL=SpecData$Cerebellum, BS=SpecData$Medulla, clade_major = SpecData$Clade_major,
                     body =   SpecData$Bodymass_g, Fossil=SpecData$Fossil, row.names=SpecData$Species )

# Use the Geometric Mean approach to size correct only
geomean <- apply(Volumes[,1:4], 1, prod)^(1/ncol(Volumes[,1:4]))#computing the geometric mean for obtaining size
vol_lsr <- cbind (log(Volumes[,1:4]/ geomean), Volumes$clade_major, geomean, Volumes$body, Volumes$Fossil) #computing the log shape ratios, 'LSR'

colnames(vol_lsr)[5] <- "clade_major"

# Reorder data to match phylogenetic tree tip label order
vol_lsr <- data.frame(vol_lsr[trees$Tree1$tip.label,]) # reorders to tip labels

PCA.vol=princomp(vol_lsr[,1:4], scores=TRUE, cor=F)


# major phylogenetic divisions

delete_singles <- c (which(vol_lsr$clade_major == "Notoryctemorphia"), which(vol_lsr$clade_major == "Paucituberculata"), which(vol_lsr$clade_major == "Stem"), which(vol_lsr$clade_major == "Dromiciops"), which(vol_lsr$clade_major == "Yalkaparidontia"))


#Doing it all in one go does not work. You have to pull out the terms first
#major_clade_vol <- procD.lm(vol_lsr[-delete_singles,1:4] ~ (droplevels(vol_lsr$clade_major[-delete_singles])))

voldelsing<-vol_lsr[-delete_singles,1:4]

cladevol <- (droplevels(as.factor(vol_lsr$clade_major[-delete_singles])))

major_clade_vol <- procD.lm(voldelsing ~ cladevol)

anova(major_clade_vol)



#### phylogenetic signal using the same function as for shape to keep things even

#the vol.lsr dataset needs special treatment before physignal will take it
vol.lsr1 <- as.matrix(vol_lsr[,1:4])

physig_vol_lsr <- matrix(NA, nrow = 3, ncol = 2)

for (i in (1:length(trees))) {
  tmp <- physignal(vol.lsr1, phy=trees[[i]])
  physig_vol_lsr[i, 1] <- tmp$phy.signal
  physig_vol_lsr[i, 2] <- tmp$pvalue
}

mean(physig_vol_lsr[, 1])
mean(physig_vol_lsr[, 2])



# Evolutionary allometry

# Brain partition volumes vs geomean
geomean_vol <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

for(i in 1:3){
  geomean_vol [[i]] <- procD.pgls(vol_lsr[,1:4]~log(vol_lsr$geomean),
                                   trees[[i]])
                                   }

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(geomean_vol)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Volumes_vs_log_Geomean"



# Brain partition volumes vs body mass
bodymass_vol <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

# I don't understand why I need to pull this out, but the code does not work otherwise

Volumes.body <- vol_lsr$Volumes.body

for(i in 1:3){
  bodymass_vol [[i]] <- procD.pgls(vol_lsr[,1:4]~log(Volumes.body),
                                   trees[[i]]
                                   )}

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(bodymass_vol)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Volumes_vs_log_Bodymass"


# how does the distribution of species in this PCA compare to the PCA of shape?
Vol.dist <- dist(PCA.vol$scores) ; Shp.dist <- dist(PCA$x)
mantel(Vol.dist,Shp.dist,method="pearson", permutations=9999)




```

#Analyses of PC2 shape variation and cerebral hemisphere volume because there seem to be some correspondence between PC2 and cerebral hemisphere partition volume

```{r}

PC2Fossil <- as.data.frame(cbind(PCA$x[,2], vol_lsr$CBR, as.factor(SpecData$Fossil)))
colnames (PC2Fossil) <- c("PC2", "CBR", "Fossil")

#if PC2 and relative  cerebral volume match, how much overall shape does relative cerebral volume explain?

#Check that species order matches up

rownames(vol_lsr) == dimnames(gdf_automatic$coords)[[3]]

#procD.pgls seems to not like the "address" for CBR so having to pull it out
CBR <- vol_lsr$CBR

Shape_vs_Cerebr <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

for(i in 1:3){
  Shape_vs_Cerebr [[i]] <- procD.pgls(gdf_automatic$coords ~ CBR,
                                   trees[[i]]
                                   )}


ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Shape_vs_Cerebr)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Cerebral_volume"




#Do Cerebral hemisphere volumes correspond to PC2?

PC2_CBR_results <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

 for(i in 1:3){
     PC2_CBR_results[[i]] <- procD.pgls(PC2 ~ CBR, trees[[i]],
                                     iter=iter, data=PC2Fossil)}


 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(PC2_CBR_results)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC2_vs_CBR_LSR"


#To further investigate the visual impression that fossils have lower PC2, get PC imporance scores

Summary_PCA <- summary(PCA)


#Who has the lowest PC2 scores - fossils are over-represented
PC2sorted <- sort(PCA$x[,2] , decreasing=FALSE)



#Do fossils have lower PC2 scores and/or cerebral volume scores?

rownames(PCA$x)==rownames (SpecData);rownames (SpecData) ==row.names(vol_lsr)

 PC2_fossil_results <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)
 CBR_vol_results <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

  for(i in 1:3){
     PC2_fossil_results[[i]] <- procD.pgls(PC2 ~ Fossil, trees[[i]],
                                     iter=iter, data=PC2Fossil)}


 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(PC2_fossil_results)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC2_in_extinct_sp."

for(i in 1:3){
     CBR_vol_results[[i]] <- procD.pgls(CBR ~ Fossil, trees[[i]],
                                     iter=iter, data=PC2Fossil)}

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(CBR_vol_results)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Cerebral_vol_in_extinct_sp."

# Similar questions surrounding Diprotodontia - do they have systematically higher PC2 or do they have significantly different shapes relative to the other species?

Dip_nonDip <- SpecData$Clade_major

# Need to manually add a level
levels(Dip_nonDip) <- c(levels(Dip_nonDip),"NonDip")
NonDip <- c(which (SpecData$Clade_major != "Diprotodontia"))
Dip_nonDip[NonDip] <- as.factor("NonDip")
Dip_nonDip <- droplevels(as.factor(Dip_nonDip))

#Add Diprotodontia to dataset, check visually afterwards that the new factor matches the species
PC2Dip <- as.data.frame(cbind(PCA$x[,2], as.factor(Dip_nonDip)), )
colnames (PC2Dip) <- c("PC2", "Dip")

 PC2_Dip_results <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

  for(i in 1:3){
     PC2_Dip_results[[i]] <- procD.pgls(PC2 ~ Dip, trees[[i]],
                                     iter=iter, data=PC2Dip)}

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(PC2_Dip_results)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC2_in_Diprotodontia"

#How about Cerebral volume in Diprotodontians?

CBR_Dip <- as.data.frame(cbind(vol_lsr$CBR, as.factor(Dip_nonDip), SpecData$Bodymass_g))

CBR_vs_Dip <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)


for(i in 1:3){
  CBR_vs_Dip [[i]] <- procD.pgls(CBR_Dip[,1]~as.factor(CBR_Dip[,2]),
                                   trees[[i]]
                                   )}

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(CBR_vs_Dip)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Cerebral_volume_in_Diprotodontians"





```

#Testing whether the cerebral volume partition or PC2 correlates with cortex volumes, using grey matter volume data from Jyothilakshmi 2020.

```{r}

#One can never check often enough
rownames(PCA$x)==rownames(vol_lsr); rownames(vol_lsr)==rownames(SpecData)

#for the cortex grey matter values from Jyothilakhsmi et al. 2020

Cortex_data <- as.data.frame(cbind(PCA$x[,2], vol_lsr$CBR, SpecData$Grey_Matter_LSR ))
colnames (Cortex_data) <- c("PC2", "CBR", "Cx_vol")
rownames(Cortex_data) <- rownames(SpecData)

#reduce down to data available
Cortex_data <- Cortex_data [which(!is.na(Cortex_data$Cx_vol)),]

#Note that P. norfolcensis and Vombatus ursinus are substantial outliers in this dataset. They have far larger/smaller cortex grey matter LSR volumes than any other marsupial.
plot(Cortex_data$CBR~Cortex_data$Cx_vol, data = Cortex_data)
text(Cortex_data$CBR~Cortex_data$Cx_vol, label = rownames(Cortex_data))


#Reduce tree to species in cortical dataset. For some reason, unlike all other analyses, this pgls throws an error if the tree is not pruned; however, I double checked with the subset of allometry without wombats and peramelemorphs and there it is fine.
omit <- which (!(rownames(SpecData) %in% rownames(Cortex_data)))


trees_Cort<- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

for(i in 1:3){
       trees_Cort[[i]] <- drop.tip(trees[[i]], omit)

     }

#Take out all species not occurring in the grey matter LSR dataset

GreyMatter_gdf <- geomorph.data.frame(coords=coords_Automatic[,,-(which (!(dimnames(coords_Automatic)[[3]] %in% rownames(Cortex_data))))], Grey_matter_lsr = Cortex_data$Cx_vol)


Shape_vs_GreyMatter_LSR <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)


for(i in 1:3){
  Shape_vs_GreyMatter_LSR [[i]] <- procD.pgls(GreyMatter_gdf$coords ~ GreyMatter_gdf$Grey_matter,
                                   trees_Cort[[i]]
                                   )}


ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Shape_vs_GreyMatter_LSR)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Grey_matter_lsr"

#Cerebral hemisphere volume relative to grey matter LSR

Cortex_results_CBR = list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

for(i in 1:3){
     Cortex_results_CBR[[i]] <- procD.pgls(Cortex_data[,2] ~ Cortex_data[,3], trees_Cort[[i]],
                                     iter=iter)
                                     }

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Cortex_results_CBR)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Cerebr_vol_vs_Grey_matter_lsr"

#Cerebral hemisphere volume relative to grey matter LSR

 Cortex_results_PC2 = list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

for(i in 1:3){
     Cortex_results_PC2[[i]] <- procD.pgls(Cortex_data[,1] ~ Cortex_data[,3], trees_Cort[[i]],
                                     iter=iter)
                                     }

 ANOVA_results_PC2 <- rbind(ANOVA_results, mean.tree.results(Cortex_results_PC2)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC2_vs_Grey_matter_lsr"





```

#Testing whether the cerebral volume partition or PC2 correlate with Neocortex volume, using data from Pirlot 1981.

```{r}

#One can never check often enough
rownames(PCA$x)==rownames(vol_lsr); rownames(vol_lsr)==rownames(SpecData)

#for the cortex grey matter values from Jyothilakhsmi et al. 2020

Neocortex_data <- as.data.frame(cbind(PCA$x[,2], vol_lsr$CBR, SpecData$Neocortex_LSR ))
colnames (Neocortex_data) <- c("PC2", "CBR", "Nc")
rownames(Neocortex_data) <- rownames(SpecData)

#reduce down to data available
Neocortex_data <- Neocortex_data [which(!is.na(Neocortex_data$Nc)),]

#Add a factor of dips and non-dips
DipFactor <- c(1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,2)
Neocortex_data <-  cbind(Neocortex_data, DipFactor)

#A quick peek
plot(Neocortex_data$CBR~Neocortex_data$Nc, data = Neocortex_data)
text(Neocortex_data$CBR~Neocortex_data$Nc, label = rownames(Neocortex_data))


#Reduce tree to species in cortical dataset. For some reason, unlike all other analyses, this pgls throws an error if the tree is not pruned; however, I double checked with the subset of allometry without wombats and peramelemorphs and there it is fine.
omit <- which (!(rownames(SpecData) %in% rownames(Neocortex_data)))


trees_NeoCort<- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

for(i in 1:3){
       trees_NeoCort[[i]] <- drop.tip(trees[[i]], omit)

     }

#Take out all species not occurring in the Neocortex LSR dataset

Neocort_gdf <- geomorph.data.frame(coords=coords_Automatic[,,-(which (!(dimnames(coords_Automatic)[[3]] %in% rownames(Neocortex_data))))], Neocort_lsr = Neocortex_data$Nc)

Shape_vs_Neocort_LSR <- list(Tree1=NULL,Tree2=NULL,Tree3=NULL)


for(i in 1:3){
  Shape_vs_Neocort_LSR [[i]] <- procD.pgls(Neocort_gdf$coords ~ Neocort_gdf$Neocort_lsr,
                                   trees_NeoCort[[i]]
                                   )}


ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Shape_vs_Neocort_LSR)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Neocortex_lsr"


Neocortex_results_CBR = list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

for(i in 1:3){
     Neocortex_results_CBR[[i]] <- procD.pgls(Neocortex_data[,2] ~ Neocortex_data[,3], trees_NeoCort[[i]],
                                     iter=iter)
                                     }

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Neocortex_results_CBR)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Cerebr_vol_vs_Neocortex_lsr"

 Neocortex_results_PC2 = list(Tree1=NULL,Tree2=NULL,Tree3=NULL)

for(i in 1:3){
     Neocortex_results_PC2[[i]] <- procD.pgls(Neocortex_data[,1] ~ Neocortex_data[,3], trees_NeoCort[[i]],
                                     iter=iter)
                                     }

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Neocortex_results_PC2)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC2_vs_Neocortex_lsr"



```




#Summarizing all analyses into one table -done throughout as well but this here works well to get a table at the end of it all.

```{r}

ANOVA_results <- data.frame( "Df"= integer(0), "F"= integer(0), "Rsq"= integer(0), "p"= integer(0))

ANOVA_results <- rbind(ANOVA_results, Csize_brainvol_output); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Csize_Brainvol"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(res.phy.allom.CS)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape vs logCsize"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(res.phy.allom.mass)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_logBodyMass"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(res.phy.allom.EQ)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape vs EQ"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(res_phy_allom_CS_VP)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape vs logCsize VP"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(res.phy.allom.mass_VP));
rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape vs log Body mass without Vombatiforms/Peramelemorphs"

ANOVA_results <-  rbind(ANOVA_results, mean.tree.results(PC1_allom_results)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC1_vs_logCsize"


ANOVA_results [length(rownames(ANOVA_results)),] <- c(

  Loco_list_Csize_inter$Tree1$aov.table[4,1],

  
 round( mean(Loco_list_Csize_inter$Tree1$aov.table$F[3],
       Loco_list_Csize_inter$Tree2$aov.table$F[3],
       Loco_list_Csize_inter$Tree3$aov.table$F[3]),
       digits=3), # average F


 round  (mean(Loco_list_Csize_inter$Tree1$aov.table$Rsq[3],
       Loco_list_Csize_inter$Tree2$aov.table$Rsq[3],
       Loco_list_Csize_inter$Tree3$aov.table$Rsq[3]), digits = 2), # average Rsq

  round( mean(Loco_list_Csize_inter$Tree1$aov.table$`Pr(>F)`[3],
       Loco_list_Csize_inter$Tree2$aov.table$`Pr(>F)`[3],
       Loco_list_Csize_inter$Tree3$aov.table$`Pr(>F)`[3]),
       digits=3)# average p
)

rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Locomotor_mode_Csize_inter"



ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Loco_list_Csize_adj)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Locomotor_mode_Csize_adj"

ANOVA_results [length(rownames(ANOVA_results)),] <- c(

   Loco_list_Csize_adj$Tree1$aov.table[4,1],

 round( mean(Loco_list_Csize_adj$Tree1$aov.table$F[2],
       Loco_list_Csize_adj$Tree2$aov.table$F[2],
       Loco_list_Csize_adj$Tree3$aov.table$F[2]),
       digits=3), # average F


 round  (mean(Loco_list_Csize_adj$Tree1$aov.table$Rsq[2],
       Loco_list_Csize_adj$Tree2$aov.table$Rsq[2],
       Loco_list_Csize_adj$Tree3$aov.table$Rsq[2]), digits = 2), # average Rsq

  round( mean(Loco_list_Csize_adj$Tree1$aov.table$`Pr(>F)`[2],
       Loco_list_Csize_adj$Tree2$aov.table$`Pr(>F)`[2],
       Loco_list_Csize_adj$Tree3$aov.table$`Pr(>F)`[2]),
       digits=3)# average p
)

rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Locomotor_mode_Csize_adj"



ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Loco_list)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape vs Locomotor mode"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Loco_list_hop)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape vs hopping"


ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Loco_list_glid)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape vs gliding"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(geomean_vol)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Volumes vs log Geomean"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(bodymass_vol)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Volumes_vs_log_Bodymass"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Shape_vs_Cerebr)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Cerebral volume"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(PC2_CBR_results)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC2_vs_CBR_LSR"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(PC2_fossil_results)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC2 in extinct sp."

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(CBR_vol_results)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Cerebral_vol_in_extinct_sp."

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(PC2_Dip_results)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC2 in Diprotodontia"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(CBR_vs_Dip)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Cerebral volume in Diprotodontia"

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Shape_vs_GreyMatter_LSR)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Grey_matter_lsr"

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Cortex_results_CBR)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Cerebr_vol_vs_Grey_matter_lsr"

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Cortex_results_PC2)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC2_vs_Grey_matter_lsr"

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Shape_vs_Neocort_LSR)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Shape_vs_Neocortex_lsr"

 ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Neocortex_results_CBR)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "Cerebr_vol_vs_Neocortex_lsr"

ANOVA_results <- rbind(ANOVA_results, mean.tree.results(Neocortex_results_PC2)); rownames(ANOVA_results)[length(rownames(ANOVA_results))] <- "PC2_vs_Neocortex_lsr"


write.csv(ANOVA_results, file = "../Figures/ANOVA_table.csv")

```






                  ####~~~~~~~ THE END ~~~~~~~####



#For figures:

```{r}
save.image("../Data/Processed/WeisbETal_Data.RData")


```